Tor's Usability for Censorship Circumvention

### 2015-12-10 paper/circumvention-experiment.tex
edit design

\noindent {\bfseries Automation}
One common piece of feedback, from experts and novices alike, was that the
connection process could be simplified with automation. At first glance, many
decisions currently made by the user could be automated. For example, bridge
selection needs to happen only if a regular connection failed, and bridges could
be tried automatically until one succeeds. However, there are security
considerations that make this solution problematic. If the launcher starts by
trying to make unobscured connections to publicly listed relays, detection is
trivial in a country that logs Internet traffic, potentially resulting in
repercussions for citizens. Furthermore, some users may be turned off by the
lack of transparency in the connection process.

However, other elements of the process can be automated without adverse effects.
For example, local proxy settings can be imported from other browsers on the
machine by looking for settings in well-known locations on the filesystem. The
same can be done to answer the question that led many of our participants
astray: is a local proxy even needed? Since this piece of automation offers a
significant potential upside, with no apparent risks, we have incorporated it
into our design. \\

\noindent {\bfseries Building a better mental model}
Participants' lack of a mental model of censorship led them to believe their
connection was censored when, in reality, it was not. It furthermore contributed
to their confusion between bridges and proxies. Our design directly addresses
these issues.

Potentially ambiguous text (i.e., referring to censorship) is removed in favor
of stating the underlying options. Separate screens with binary options (Do you
need a proxy? Yes/No) are removed; instead the option to opt out is incorporated
on the configuration screen. This results in a more streamlined interface (fewer
clicks required) and unified the context of each option: all information about
bridges (or proxies) is now contained on one screen, so the user has the full
context of their decision. \\

@@

A clear example of this is the ``Connect'' button on the first screen, which has
increased in size and become green, making it more prominent as the recommended
option. A similar aesthetic applies to the connect button seen after custom
configuration.

Rarely used elements are deemphasized. For example, the custom bridges box,
which previously took up a significant portion of the bridge configuration
screen, is now hidden until that option is selected.

A progress indicator at the top of the window indicates which configuration step
the user is on. It serves to clarify the independence of bridges and proxies,
which need to be configured separately. In the event of a failure, the user is
returned to the first step of the process -- the bridge configuration -- to
avoid misleading users into thinking the proxy settings are the culprit. \\

@@

Many users were led astray by interface elements that were unhelpful or
misleading if the program encountered connection problems. Our design seeks to
prevent such confusion by gracefully handling failures.

The progress bar in the connection window drew many complaints, because it got
stuck and didn't offer any guidance for how long users were supposed to wait. We
address this by providing a countdown timer, indicating how long users should
expect to wait. We further enhance the progress bar by adding visual indicators
for major milestones in the connection, providing better feedback about what has
been accomplished so far. Additionally, the current status of the connection
attempt (drawn from the log) is displayed, to aid in understanding and
debugging.

In the event that the connection attempt fails, messages from the log are
displayed, along with tips and suggestions about what to do next (e.g., try
another bridge). To make potential the options clearer, the new failure screen
includes a separate button for retrying without changing settings and another
for returning to reconfigure the application. \\

\noindent {\bfseries Additional changes}
In addition to the changes above, we have simplified flows where possible,
provided clearer definitions of bridges and proxies, and reduced visual clutter.
{\color {red} Anything else?}

>>>>

\noindent {\bfseries Automation}
One common piece of feedback from experts and novices alike was that the
connection process could be simplified with automation. At first glance, many
decisions currently made by the user could be automated. For example, bridge
selection needs to happen only if a regular connection failed, and bridges could
be tried automatically until one succeeds. However, there are security
considerations that make this solution problematic. If the launcher starts by
trying to make unobscured connections to publicly listed relays, detection is
trivial in a country that logs Internet traffic, potentially resulting in
repercussions for citizens. Furthermore, users may be turned off by the
lack of transparency in the connection process.

However, other elements of the process can be automated without adverse effects.
For example, local proxy settings can be imported from other browsers on the
machine by looking for settings in well-known locations on the filesystem. Automatically
detecting the need for a proxy and importing the proxy settings required would save
our participants from a task they do not have the background to easily complete. 
Since this piece of automation offers a significant increase in usability for our users
(addressing our most common failure case) with no apparent risks, we have incorporated it
into our design. \\

\noindent {\bfseries Building a better mental model}
Participants' lack of a mental model of censorship led them to believe their
Tor connection was censored when, in reality, it was not. It furthermore contributed
to their confusion between bridges and proxies. Our design directly addresses
these issues.

Potentially ambiguous text (i.e., referring to censorship) is removed in favor
of stating the underlying options, as it was shown to confuse our participants.
Previously, separate screens with binary options (Do you need a proxy? Yes/No) 
were provided to help users decide if they needed a bridge or proxy, and skipping
the configuration windows for bridges and proxies, respectively, when users
answered that they didn't need one. We have removed these screens
because our participants did not understand how to answer them correctly. This is 
anticipated in the old interface, in fact: written text in the interface guides users on to 
how to answer those questions.

We chose to display all information about bridges and proxies is now contained on one 
respective screen in the interface, giving the user has the full context of their decision.
We made the guidance from the last interface (to not choose a bridge, to not choose a proxy)
more direct by selecting the  option to opt out of a proxy as the default option on those screens. 
This results in a more streamlined interface with fewer windows
and fewer clicks required for our participants. \\

@@

A clear example of this is the ``Connect'' button on the first screen, which has
increased in size and become green, making it more prominent as the recommended
option. We believe that this will help the majority of Tor users. 
Although some of our participants in environments had chosen to connect
directly when they needed to configure, this is generally seen as low-risk today, and 
the majority of censorship today can be circumvented with a direct connection. 
A similar aesthetic applies to the connect button seen after custom
configuration.

To help lessen the cognitive load and distribute a representative amount of space
for options, recommendations are selected as the default option, and rarely used 
elements are deemphasized. To encourage our users to click on helpful options, 
the help button on the bridge configuration screen
has been emphasized to be bigger, since it is a helpful option which people rarely
clicked. To discourage our users from using an advanced setting only technical users 
have a chance of  configuring correctly and should only be used in very rare circumstances, 
we have now incorporated the custom bridge configuration in the dropdown menu; 
it previously took up a significant portion of the bridge configuration screen.

A progress indicator at the top of the window indicates which configuration step
the user is on. It serves to clarify the independence of bridges and proxies,
which need to be configured separately. In the event of a failure, the user is
returned to the first step of the process -- the bridge configuration -- to
avoid misleading users into thinking the proxy settings are the culprit. Previously,
the user had to manually click to go backward in the dialog to configure bridges.
Even in the rare event that the proxy is the issue for the connection setting, the user
will be directed to the proxy page naturally through the flow of the UI, without
having the user to defy the natural flow of configuration. \\

@@

Many users were led astray by interface elements that were unhelpful or
misleading if the program encountered connection problems. Our design seeks to
prevent such confusion by anticipating failures and providing specific instructions
for our participants on their next steps.

The progress bar in the connection window drew many complaints, because it got
stuck and didn't offer any guidance for how long users were supposed to wait. 
The connection process can take a non-trivial couple of minutes, and users may
think that the connection isn't working, even when it is making slow progress. We
address this by providing a countdown timer, indicating that a long wait time does
not indicate failure. We further encourage waiting by adding visual indicators
for major milestones in the connection, providing better feedback about what has
been accomplished so far. Additionally, the current status of the connection
attempt (drawn from the log) is displayed, to aid in understanding and
debugging.

In the event that the connection attempt fails, messages from the log are
displayed, along with tips and suggestions about what to do next (e.g., try
another bridge). To make potential the options clearer, the new failure screen
includes a separate button for common next steps which are able to resolve
a failed connection, which are to retry without changing settings and to reconfigure 
the connection. \\

\noindent {\bfseries Additional changes}
In addition to the changes above, we have simplified flows where possible,
provided clearer definitions of bridges and proxies, and reduced visual clutter.
We believe that all of these changes are cohesive with the other design changes
made to the interface and will be beneficial for the user.
